### 2021-01-07 paper/natastron/2nd_draft/paper.tex
Ready to circulate

Gyrochronology was previously calibrated on stellar clusters, which have well constrained ages, but only up to roughly $4\, \rm{Gyr}$ \cite{meibom+2015, barnes+2016}. Recently, ages of main sequence field stars (i.e. not in clusters) observed by \kepler \cite{borucki+2010} became more widely available through asteroseismology, the study of stellar variability \cite{silvaaguirre+2015}. When looking at these stars, disagreements were found between the observations and gyrochronology beyond the middle of stars' main sequence lifetimes, which could not be reconciled with existing theories \cite{angus+2015, nielsen+2015, davies+2015}. It was proposed that at some stage in a star's evolution it undergoes \textit{weakened magnetic braking} (WMB, \cite{vansaders+2016}), where the efficiency of angular momentum loss rapidly drops. This would cause stars to maintain relatively fast rotation rates at old ages, which we would not expect from the extrapolation of existing gyrochronology relations.

@@

Asteroseismic rotation requires a faster observing cadence and higher signal-to-noise than spot rotation, but does not require spots to be visible, meaning that we can measure rotation for older, quiescent stars that would not have been present in existing rotation catalogues. 
%This makes asteroseismic rotation rates a perfect test of weakened magnetic braking on the main sequence. 
Here, we measured asteroseismic rotation rates of 91 main sequence stars across a broad range of  ages, and determined whether this new ensemble agrees more closely with a classical rotational evolution scenario, or one in which weakened magnetic braking takes place.\\

@@

We compared our results with both published and unpublished asteroseismic rotation rates \cite{nielsen+2015,davies+2015,davies+2016, lund+2017, benomar+2018}, and found that the three stars with the lowest angles of inclination ($< 10^\circ$) were inconsistent with multiple other studies. To some degree this was expected, as  for stars viewed near pole-on the split modes of oscillation have extremely low power, making them hard to detect reliably. We flagged our rotation rates for these stars, and did not use them in the next steps in our analysis, leaving a sample of 91 stars for which we considered our asteroseismic rotation measurements to be robust (for a full justification, see \textbf{Methods}).\\

@@

Figure \ref{fig:protlit} shows a comparison between 48 stars from our sample for which spot  rotation rates had previously been measured. Rotation rates from spot modulation are subject to measuring multiples of the true rotation rate, and also sometimes over-correct for this effect. This is not the case for asteroseismic rotation, and so some stars may appear at the 2:1 and 1:2 lines in the figure. We fit a line of the form $P_{\rm surf} = m \times P_{\rm seis}$, using the larger of the asymmetrical uncertainties on asteroseismic rotation, and excluding stars above the 1.8:1 line to avoid likely multiples of the true rotation rate acting as outliers (transparent in the Figure, see \textit{\textbf{Methods}} for details.). Our fit found a value of $m = 0.96 \pm 0.03$, showing a close agreement ($<2\sigma$) between asteroseismic and surface rotation rates on a population level. =
%Of the stars used in this comparison, 21 had a median value of $Q_{\rm WMB}$ in the gyrochronology analysis. We repeated the model fit for these stars only, and found a value of $m = 0.96 \pm 0.04$.  
This overall agreement between spot and seismic rotation rates is in line with previous studies of Sun-like stars \cite{nielsen+2015, benomar+2015, gizon+2013, chaplin+2013}, and so we concluded that our asteroseismic ensemble can be reliably used to draw inferences about gyrochronology.\\

\begin{figure}[h!]
	\centering
	\includegraphics[width=0.6\textwidth]{Images/surf-seis-comparison_update.pdf}
	\caption{Comparisons between asteroseismic and photometric measures of stellar rotation, for 48 stars. Literature values were taken, in order of priority, from: Garcia et al. (2014) \cite{garcia+2014} and Nielsen et al. (2013) \cite{nielsen+2013} (i.e. if a value was reported in Nielsen et al. (2013) and also in Garcia et al. (2014), the latter was used). Stars used in the original van Saders et al. (2019) \cite{vansaders+2016} study that first proposed weakened magnetic braking are highlighted. The four triangles represent stars included in the Nielsen et al. (2015) study \cite{nielsen+2015} that previously measured this relationship. The solid line indicates 1:1 line, while the dash-dotted lines represent 2:1 and 1:2 lines. The shaded region around the 1:1 line is the $1\sigma$ credible interval of the fit to the data, the result of which is shown as the title of the figure. Stars that were not included in the fitting process are transparent. The errorbars on the asteroseismic rotation represent the $15.9^{\rm th}$ and $84.1^{\rm st}$ percentiles on the measured posterior distributions, and are smaller than the points in some cases.}
	\label{fig:protlit}
\end{figure}

\textbf{Comparing gyrochronology models:} To evaluate the implications of our seismic ensemble for gyrochronology, we compared our new asteroseismic rotation rates to two population models of rotational evolution \cite{vansaders+2019}: a `standard' model, which assumes a traditional angular momentum transport through magnetically driven stellar winds \cite{vansaders+2016,skumanich1972, kawaler1988}, and one where weakened magnetic braking takes place (hereafter the `WMB model') \cite{vansaders+2016}.  The WMB model was identical to the standard model in its input physics, except for the condition that angular momentum loss ceases above a critical Rossby number of $Ro_{\rm crit} = 1.97$\footnote{The Rossby number may be defined as $Ro \equiv P/\tau_{\rm cz}$, where $\tau_{\rm cz}$ is the convective turnover timescale, and $P$ is the rotation. The Rossby number roughly scales inversely with stellar activity (and so increases as stars age).}.

Instead of comparing our rotation rates directly to models of individual stars, we instead compared them to synthetic TRILEGAL populations based on what the \textit{Kepler} field would look like under these two different braking laws \cite{vansaders+2019, girardi+2012}. These synthetic populations were then altered to match the distribution of existing catalogues of stars in the \kepler field in order to replicate the \kepler selection effects \cite{berger+2020}, and evolved following a standard or WMB braking law. For the full details on how these models were constructed, see \textbf{Methods}.

@@

We evaluated our ensemble against both population models simultaneously using a Bayesian mixture model, which treated the data as being drawn from a mixture made from Kernel Density Estimates (KDEs) of both populations, modulated by a mixing parameter $Q_{\rm WMB}$. In the limit that $Q_{\rm WMB} \rightarrow 1$, the data are most likely drawn from the WMB model. In the limit $Q_{\rm WMB} \rightarrow 0$, the standard model is more likely. This mixture model was evaluated in a five-dimensional parameter space, accounting for the population distributions in effective temperature ($T_{\rm eff}$), metallicity (\feh), asteroseismic mass ($M$) and age ($t$) and our new rotation periods ($P$). For further detail, see \textbf{Methods}.\\

@@

Both hot and sub-giant stars do not strongly prefer one model over the other (with both cumulative probabilities crossing the $Q_{\rm{WMB}} = 0.5$ line at roughly $60\%$), which is expected given that these stars in our sample all lie at hot temperatures near the Kraft break. When performing this analysis on all 89 stars for which seismic rotation was measured (excluding the two stars with low metallicity), the total probability above $Q_{\rm{WMB}} = 0.5$ is 96.6\%. \rtwo{The total posterior probability for all stars above $0.6$ and $0.7$ is $93.3\%$ and $78.2\%$  respectively. This does not necessarily mean that a mixture of the two models is preferred, as stars that have not yet undergone weakened braking, or which are near the transition period, will have a flat posterior distribution with no strong preference for either model.}\\

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{modelresults.pdf}
	\caption{Posterior estimates of the mixture model parameter $Q_{\rm WMB}$, broken down by stellar classification. A value of $Q_{\rm WMB}$ close to 1 indicates that the data are more consistent with a rotational evolution that includes weakened magnetic braking, whereas a value close to 0 indicates that the data are more consistent with a standard rotational evolution scenario. \textit{Left:} the product of histograms of the posterior estimates for 73 stars, or subsets thereof for different stellar classifications. \textit{Right:} the cumulative posterior probability for 73 stars, or subsets thereof for different stellar types. On the y-axis of the left hand plot, $p(Q_{\rm WMB} | \mathcal{D})$ represents the posterior distribution on $Q_{\rm WMB}$ given the data. The product symbol represents the sum of the posteriors for individual stars $i$, for a total of $N$ stars (which differs depending on the stellar type being shown).}
	\label{fig:gyroresults}
\end{figure}

\textbf{Verifying consequences for gyrochronology:} In order to verify that our results presented a meaningful confirmation of weakened magnetic braking, we ensured that the same results were obtained when using different subsets of the ensemble, and when exploring potential biases. We found that the joint posterior distribution did not change significantly when excluding stars with metallicities that fell only slightly outside the model range, or those with extremely small uncertainties on asteroseismic mass (see \textbf{Methods}). Stars in our sample with known binary companions that may disrupt their rotational evolution were not found to contribute strongly to the joint posterior probability of $Q_{\rm WMB}$.

@@

A recent study using spectroscopic rotation measurements of 14 solar twins, compared to grids of stellar models, found marginal evidence favouring a standard rotational evolution \cite{lorenzo-oliveira+2019}. If weakened magnetic braking were to take place, this work proposed that it would most likely be outside the range of their sample, at $Ro_{\rm {crit}} \geq 2.29$, compared to the value of $1.97$ used in this work. Our sample does not overlap with theirs, so we can not directly compare to their results. Our work does find statistical agreement with a model of weakened magnetic braking at a critical Rossby number of $1.97$ over one without any braking at all, indicating that traces of weakened magnetic braking do occur at Rossby numbers of around $2$. A comparison to models of different Rossby numbers will be explored in future papers, using the LEGACY and `Kages' rotation sample presented in this work.

As a sanity check that the evidence in favour of weakened magnetic braking presented in this work is independently robust, we recreated the joint posterior probability for $Q_{\rm WMB}$, this time looking only at MS stars, and excluding any of the 22 stars that were included in the original van Saders study used to detect weakened magnetic braking. Of the 47 MS stars used to distinguish between stellar models, 16 were included in the van Saders sample. Both the 16 stars in the original van Saders study as well as the remaining MS stars favoured the WMB model, although the van Saders stars did so more strongly, not including any stars that favoured the standard model. Specifically, when considering only the van Saders stars, $96.3\%$ of the total joint probability lay above $Q_{\rm WMB} = 0.5$, compared to $91.7\%$ when considering only those stars \emph{not} included in the van Saders study. The stars initially used to propose the WMB model are still those most strongly in favour of it even when using asteroseismic rotation rates only, a result which is now further reinforced by an expanded sample that includes quiescent stars.\\

\textbf{Conclusions:} In this work, we obtained asteroseismic rotation rates for 91 stars using a hierarchical Bayesian model to fit for oscillation frequencies, convective background, and rotational splitting simultaneously. We validated these results against previously published and unpublished asteroseismic rotation measurements, making our new rotation catalogue the largest self-consistent sample of asteroseismic rotation in main sequence stars to date. We evaluated the stars in our sample against synthetic population models of the \textit{Kepler} field evolved under either a `standard' model of rotational evolution, or one where weakened magnetic braking takes place. This evaluation was done using a Bayesian mixture model to determine whether the ensemble preferred one model over another, and was done simultaneously in asteroseismic rotation, mass, age, and spectroscopic temperature and metallicity.\\

We leave the reader with the following conclusions:
\begin{enumerate}
	\item We found that our ensemble was $98.4\%$ more likely to be drawn from a weakened magnetic braking model than a standard model of rotational evolution \cite{vansaders+pinsonneault2013}, at a critical Rossby number of $Ro_{\rm{crit}} = 1.97$. This work expands upon previous analysis \cite{vansaders+2019} by including quiescent stars older than the Sun. This conclusion was found to be robust against change of ensemble members and potentially underestimated asteroseismic systematic uncertainties. A comparison to other braking laws \cite{matt+2015} and other gyrochronology relations \cite{barnes2010} using our asteroseismic ensemble will take place in future work.

>>>>

Gyrochronology was previously calibrated on stellar clusters, which have well constrained ages, but only up to roughly $4\, \rm{Gyr}$ \cite{meibom+2015, barnes+2016}. Recently, ages of main sequence field stars (i.e. not in clusters) observed by \kepler \cite{borucki+2010} became more widely available through asteroseismology, the study of stellar variability \cite{silvaaguirre+2015}. When looking at these stars, disagreements were found between observations and gyrochronology beyond the middle of stars' main sequence lifetimes, which could not be reconciled with existing theories \cite{angus+2015, nielsen+2015, davies+2015}. It was proposed that at some stage in a star's evolution it undergoes \textit{weakened magnetic braking} (WMB, \cite{vansaders+2016}), where the efficiency of angular momentum loss rapidly drops. This would cause stars to maintain relatively fast rotation rates at old ages, which we would not expect from the extrapolation of existing gyrochronology relations.

@@

Asteroseismic rotation requires a faster observing cadence and higher signal-to-noise than spot rotation, but does not require spots to be visible, meaning that we can measure rotation for older, quiescent stars that would not have been present in existing rotation catalogues. 
%This makes asteroseismic rotation rates a perfect test of weakened magnetic braking on the main sequence. 
Here, we measured asteroseismic rotation rates of 91 main sequence stars across a broad range of  ages, and determined whether \rtwo{these new data agree} more closely with a classical rotational evolution scenario, or one in which weakened magnetic braking takes place.\\

@@

We compared our results with both published and unpublished asteroseismic rotation rates \cite{nielsen+2015,davies+2015,davies+2016, lund+2017, benomar+2018}, and found that the three stars with the lowest angles of inclination ($< 10^\circ$) were inconsistent with multiple other studies. For stars viewed near pole-on the split modes of oscillation have extremely low power, making them hard to detect reliably \cite{lund+2014}. We flagged our rotation rates for these stars, and did not use them in the next steps in our analysis, leaving a sample of 91 stars for which we considered our asteroseismic rotation measurements to be robust (for a full justification, see \textbf{Methods}).\\

@@

Figure \ref{fig:protlit} shows a comparison between 48 stars from our sample for which spot  rotation rates had previously been measured. Rotation rates from spot modulation are subject to measuring multiples of the true rotation rate, and also sometimes over-correct for this effect. This is not the case for asteroseismic rotation, and so some stars may appear at the 2:1 and 1:2 lines in the figure. We fit a line of the form $P_{\rm surf} = m \times P_{\rm seis}$, using the larger of the asymmetrical uncertainties on asteroseismic rotation, and excluding stars above the 1.8:1 line to avoid likely multiples of the true rotation rate acting as outliers (transparent in the Figure, see \textbf{Methods}å for details). Our fit found a value of $m = 0.96 \pm 0.03$, showing a close agreement ($<2\sigma$) between asteroseismic and surface rotation rates on a population level. This overall agreement between spot and seismic rotation rates is in line with previous studies of Sun-like stars \cite{nielsen+2015, benomar+2015, gizon+2013, chaplin+2013}, and so we concluded that our asteroseismic ensemble can be reliably used to draw inferences about gyrochronology.\\

\begin{figure}[h!]
	\centering
	\includegraphics[width=0.6\textwidth]{Images/surf-seis-comparison_update.pdf}
	\caption{Comparisons between asteroseismic and photometric measures of stellar rotation, for 48 stars. Literature values were taken Garcia et al. (2014) \cite{garcia+2014} and Nielsen et al. (2013) \cite{nielsen+2013} (if a value was reported in both, the Garcia et al. 2014 value was used). Stars used in the original van Saders et al. (2019) \cite{vansaders+2016} study that first proposed weakened magnetic braking are highlighted. The four triangles represent stars included in the Nielsen et al. (2015) study \cite{nielsen+2015} that previously measured this relationship. The solid line indicates 1:1 line, while the dash-dotted lines represent 2:1 and 1:2 lines. The shaded region around the 1:1 line is the $1\sigma$ credible interval of the fit to the data, the result of which is shown as the title of the figure. Stars that were not included in the fitting process are transparent. The errorbars on the asteroseismic rotation represent the $15.9^{\rm th}$ and $84.1^{\rm st}$ percentiles on the measured posterior distributions, and are smaller than the points in some cases.}
	\label{fig:protlit}
\end{figure}

\textbf{Comparing gyrochronology models:} To evaluate the implications of our \rtwo{new seismic rotation rates} for gyrochronology, we compared \rtwo{the stars in our ensemble} to two population models of rotational evolution \cite{vansaders+2019}: a `standard' model, which assumes a traditional angular momentum transport through magnetically driven stellar winds \cite{vansaders+2016,skumanich1972, kawaler1988}, and one where weakened magnetic braking takes place (hereafter the `WMB model') \cite{vansaders+2016}.  The WMB model was identical to the standard model in its input physics, except for the condition that angular momentum loss ceases above a critical Rossby number of $Ro_{\rm crit} = 1.97$\footnote{The Rossby number may be defined as $Ro \equiv P/\tau_{\rm cz}$, where $\tau_{\rm cz}$ is the convective turnover timescale, and $P$ is the rotation. The Rossby number roughly scales inversely with stellar activity (and so increases as stars age).}. \rtwo{Because of this, the distribution of stars with a Rossby number below this threshold will be identical in both models}.

Instead of comparing our rotation rates directly to models of individual stars, we instead compared them to \rtwo{the distribution of} synthetic TRILEGAL populations based on what the \textit{Kepler} field would look like under these two different braking laws \cite{vansaders+2019, girardi+2012}. These synthetic populations were then altered to match the distribution of existing catalogues of stars in the \kepler field in order to replicate the \kepler selection effects \cite{berger+2020}, and evolved following a standard or WMB braking law. For details on how these models were constructed, see \textbf{Methods}.

@@

We evaluated our ensemble against both population models simultaneously using a Bayesian mixture model, which treated the data as being drawn from a mixture made from Kernel Density Estimates (KDEs) of both populations, modulated by a mixing parameter $Q_{\rm WMB}$. In the limit that $Q_{\rm WMB} \rightarrow 1$, the data are most likely drawn from the WMB model. In the limit $Q_{\rm WMB} \rightarrow 0$, the standard model is more likely. This mixture model was evaluated in a five-dimensional parameter space, accounting for the population distributions in effective temperature ($T_{\rm eff}$), metallicity (\feh), asteroseismic mass ($M$) and age ($t$) and our new rotation periods ($P$). For further details, see \textbf{Methods}.\\

@@

Both hot and sub-giant stars do not strongly prefer one model over the other (with both cumulative probabilities crossing the $Q_{\rm{WMB}} = 0.5$ line at roughly $60\%$). \rtwo{This indicates that these stars lie below our model's critical Rossby number (and therefore in the parameter space where our two models are identical), or that they are only weakly affected by angular momentum loss (as is the case for hot stars near the Kraft break).} When performing this analysis on all 89 stars for which seismic rotation was measured (excluding the two stars with low metallicity), the total probability above $Q_{\rm{WMB}} = 0.5$ is 96.6\%. \rtwo{The total posterior probability for all stars above $Q_{\rm{WMB}} = 0.6$ and $0.7$ is $93.3\%$ and $78.2\%$  respectively. This does not necessarily mean that a mixture of the two models is preferred, as stars that have not yet undergone weakened braking, or which are near the transition period, will have a flat posterior distribution with no strong preference for either model.}\\

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{modelresults.pdf}
	\caption{Posterior estimates of the mixture model parameter $Q_{\rm WMB}$, broken down by stellar classification. A value of $Q_{\rm WMB}$ close to 1 indicates that the data are more consistent with a rotational evolution that includes weakened magnetic braking, whereas a value close to 0 indicates that the data are more consistent with a standard rotational evolution scenario. \textit{Left:} \rtwo{the joint posterior probabilities} for 73 stars, or subsets thereof for different stellar classifications. \textit{Right:} the cumulative posterior probability for 73 stars, or subsets thereof for different stellar types. On the y-axis of the left hand plot, $p(Q_{\rm WMB} | \mathcal{D})$ represents the posterior distribution of $Q_{\rm WMB}$ given the data. The product symbol represents the combination of the posteriors for individual stars $i$, for a total of $N$ stars (which differs depending on the stellar type being shown).}
	\label{fig:gyroresults}
\end{figure}

\textbf{Verifying consequences for gyrochronology:} In order to verify that our results presented a meaningful \rtwo{conclusion about} weakened magnetic braking, we ensured that the same results were obtained when using different subsets of the ensemble, and when exploring potential biases. We found that the joint posterior distribution did not change significantly when excluding stars with metallicities that fell only slightly outside the model range, or those with extremely small uncertainties on asteroseismic mass (see \textbf{Methods}). Stars in our sample with known binary companions that may disrupt their rotational evolution were not found to contribute strongly to the joint posterior probability of $Q_{\rm WMB}$.

@@

A recent study using spectroscopic rotation measurements of 14 solar twins, compared to grids of stellar models, found marginal evidence favouring a standard rotational evolution \cite{lorenzo-oliveira+2019}. If weakened magnetic braking were to take place, this work proposed that it would most likely be outside the range of their sample, at $Ro_{\rm {crit}} \geq 2.29$, compared to the value of $1.97$ used in this work. Our sample does not overlap with theirs, so we can not directly compare to their results.
The statistical agreement between our seismic stars and a model of weakened magnetic braking at a critical Rossby number of $1.97$ over the standard scenario indicates that traces of weakened magnetic braking do occur at Rossby numbers of around $2$. \rtwo{However, we only intend to show that asteroseismic rotation rates indicate a preference for the presence of weakened magnetic braking, instead of claiming a particular critical Rossby number is correct. A measurement of the exact critical Rossby number at which these occurs, using individual models of our seismic targets, is the subject of a paper in preparation.}

As a sanity check that the evidence in favour of weakened magnetic braking presented in this work is independently robust, we recreated the joint posterior probability for $Q_{\rm WMB}$, this time looking only at MS stars, and excluding any of the 22 stars that were part of the original van Saders study used to propose weakened magnetic braking. Of the 47 MS stars used to distinguish between stellar models, 16 were included in the van Saders sample. Both the 16 stars in the original van Saders study as well as the remaining MS stars favoured the WMB model, although the van Saders stars did so more strongly, not including any stars that favoured the standard model. Specifically, when considering only the van Saders stars, $96.3\%$ of the total joint probability lay above $Q_{\rm WMB} = 0.5$, compared to $91.7\%$ when considering only those stars \emph{not} included in the van Saders study. The stars initially used to propose the WMB model are still those most strongly in favour of it even when using asteroseismic rotation rates only, a result which is now further reinforced by an expanded sample that includes quiescent stars.\\

\textbf{Conclusions:} In this work, we obtained asteroseismic rotation rates for 91 stars using a hierarchical Bayesian model to fit for oscillation frequencies, convective background, and rotational splitting simultaneously. We validated these results against previously published and unpublished asteroseismic rotation measurements, making our new rotation catalogue the largest self-consistent sample of asteroseismic rotation in main sequence stars to date. We evaluated the stars in our sample against synthetic population \rtwo{distribution} models of the \textit{Kepler} field evolved under either a `standard' model of rotational evolution, or one where weakened magnetic braking takes place. This evaluation was done using a Bayesian mixture model to determine whether the ensemble preferred one model over another, and was done simultaneously in asteroseismic rotation, mass, age, and spectroscopic temperature and metallicity.\\

We leave the reader with the following conclusions:
\begin{enumerate}
	\item We found that our ensemble was $98.4\%$ more likely to be drawn from a weakened magnetic braking model than a standard model of rotational evolution \cite{vansaders+pinsonneault2013}, at a critical Rossby number of $Ro_{\rm{crit}} = 1.97$. \rtwo{Stars in our ensemble overall show a weaker rate of braking than one would expect from solar-calibrated rotational evolution models.} This work expands upon previous analysis \cite{vansaders+2019} by including quiescent stars older than the Sun. This conclusion was found to be robust against change of ensemble members and potentially underestimated asteroseismic systematic uncertainties. A comparison to other braking laws \cite{matt+2015} and other gyrochronology relations \cite{barnes2010} using our asteroseismic ensemble will take place in future work.
